### 2021-03-25 paper/main.tex
added TEX file for final paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%2345678901234567890123456789012345678901234567890123456789012345678901234567890
%        1         2         3         4         5         6         7         8

\documentclass[letterpaper, 10 pt, conference]{ieeeconf}  % Comment this line out
                                                          % if you need a4paper
%\documentclass[a4paper, 10pt, conference]{ieeeconf}      % Use this line for a4
                                                          % paper

\IEEEoverridecommandlockouts                              % This command is only
                                                          % needed if you want to
                                                          % use the \thanks command
\overrideIEEEmargins
% See the \addtolength command later in the file to balance the column lengths
% on the last page of the document

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

% The following packages can be found on http:\\www.ctan.org
%\usepackage{graphics} % for pdf, bitmapped graphics files
%\usepackage{epsfig} % for postscript graphics files
%\usepackage{mathptmx} % assumes new font selection scheme installed
%\usepackage{mathptmx} % assumes new font selection scheme installed
%\usepackage{amsmath} % assumes amsmath package installed
%\usepackage{amssymb}  % assumes amsmath package installed

\title{\LARGE \bf
Tibetan Machine Translation 
}

%\author{ \parbox{3 in}{\centering Huibert Kwakernaak*
%         \thanks{*Use the $\backslash$thanks command to put information here}\\
%         Faculty of Electrical Engineering, Mathematics and Computer Science\\
%         University of Twente\\
%         7500 AE Enschede, The Netherlands\\
%         {\tt\small h.kwakernaak@autsubmit.com}}
%         \hspace*{ 0.5 in}
%         \parbox{3 in}{ \centering Pradeep Misra**
%         \thanks{**The footnote marks may be inserted manually}\\
%        Department of Electrical Engineering \\
%         Wright State University\\
%         Dayton, OH 45435, USA\\
%         {\tt\small pmisra@cs.wright.edu}}
%}

\author{ Jacob Atkins$^{1}$, Chris Calloway$^{2}$, Kelsey Namba$^{3}$, Rahul Sehrawat$^{4}$% <-this % stops a space
\thanks{*Smara Muresan, Isabelle Zuagg, Jonathan Reeve}% <-this % stops a space

}

\begin{document}

\maketitle
\thispagestyle{empty}
\pagestyle{empty}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}

(Rahul)

\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{INTRODUCTION}

(Kelsey)

\section{PROCEDURE}

\subsection{Preprocessing}

(Chris)

\subsection{Model Design}

(Jacob)

\section{Data}

(not sure if we need this section, maybe can be merged with another)

\section{Results}

(TODO)

\section{Conclusion}

(TODO)

\section*{APPENDIX}

Appendixes should appear before the acknowledgment. (not sure what we even need to put here)

\section*{ACKNOWLEDGMENT}

(Thanks to professors + COMS 4995, etc etc)

(Put in all papers from lit review here)

(Chris)
(Rahul)
(Kelsey)
\begin{thebibliography}{99}

\bibitem{c1} G. O. Young, ``Synthetic structure of industrial plastics (Book style with paper title and editor),'' 	in Plastics, 2nd ed. vol. 3, J. Peters, Ed.  New York: McGraw-Hill, 1964, pp. 15--64.
\bibitem{c2} W.-K. Chen, Linear Networks and Systems (Book style).	Belmont, CA: Wadsworth, 1993, pp. 123--135.
\bibitem{c3} H. Poor, An Introduction to Signal Detection and Estimation.   New York: Springer-Verlag, 1985, ch. 4.
\bibitem{c4} B. Smith, ``An approach to graphs of linear forms (Unpublished work style),'' unpublished.
\bibitem{c5} E. H. Miller, ``A note on reflector arrays (Periodical styleÑAccepted for publication),'' IEEE Trans. Antennas Propagat., to be publised.
\bibitem{c6} J. Wang, ``Fundamentals of erbium-doped fiber amplifiers arrays (Periodical styleÑSubmitted for publication),'' IEEE J. Quantum Electron., submitted for publication.
\bibitem{c7} C. J. Kaufman, Rocky Mountain Research Lab., Boulder, CO, private communication, May 1995.
\bibitem{c8} Y. Yorozu, M. Hirano, K. Oka, and Y. Tagawa, ``Electron spectroscopy studies on magneto-optical media and plastic substrate interfaces(Translation Journals style),'' IEEE Transl. J. Magn.Jpn., vol. 2, Aug. 1987, pp. 740--741 [Dig. 9th Annu. Conf. Magnetics Japan, 1982, p. 301].
\bibitem{c9} M. Young, The Techincal Writers Handbook.  Mill Valley, CA: University Science, 1989.
\bibitem{c10} J. U. Duncombe, ``Infrared navigationÑPart I: An assessment of feasibility (Periodical style),'' IEEE Trans. Electron Devices, vol. ED-11, pp. 34--39, Jan. 1959.
\bibitem{c11} S. Chen, B. Mulgrew, and P. M. Grant, ``A clustering technique for digital communications channel equalization using radial basis function networks,'' IEEE Trans. Neural Networks, vol. 4, pp. 570--578, July 1993.
\bibitem{c12} R. W. Lucky, ``Automatic equalization for digital communication,'' Bell Syst. Tech. J., vol. 44, no. 4, pp. 547--588, Apr. 1965.
\bibitem{c13} S. P. Bingulac, ``On the compatibility of adaptive controllers (Published Conference Proceedings style),'' in Proc. 4th Annu. Allerton Conf. Circuits and Systems Theory, New York, 1994, pp. 8--16.
\bibitem{c14} G. R. Faulhaber, ``Design of service systems with priority reservation,'' in Conf. Rec. 1995 IEEE Int. Conf. Communications, pp. 3--8.
\bibitem{c15} W. D. Doyle, ``Magnetization reversal in films with biaxial anisotropy,'' in 1987 Proc. INTERMAG Conf., pp. 2.2-1--2.2-6.
\bibitem{c16} G. W. Juette and L. E. Zeffanella, ``Radio noise currents n short sections on bundle conductors (Presented Conference Paper style),'' presented at the IEEE Summer power Meeting, Dallas, TX, June 22--27, 1990, Paper 90 SM 690-0 PWRS.
\bibitem{c17} J. G. Kreifeldt, ``An analysis of surface-detected EMG as an amplitude-modulated noise,'' presented at the 1989 Int. Conf. Medicine and Biological Engineering, Chicago, IL.
\bibitem{c18} J. Williams, ``Narrow-band analyzer (Thesis or Dissertation style),'' Ph.D. dissertation, Dept. Elect. Eng., Harvard Univ., Cambridge, MA, 1993. 
\bibitem{c19} N. Kawasaki, ``Parametric study of thermal and chemical nonequilibrium nozzle flow,'' M.S. thesis, Dept. Electron. Eng., Osaka Univ., Osaka, Japan, 1993.
\bibitem{c20} J. P. Wilkinson, ``Nonlinear resonant circuit devices (Patent style),'' U.S. Patent 3 624 12, July 16, 1990.

\end{thebibliography}

\end{document}

### 2021-03-25 paper/main.tex
added a little bit of procedure

@@
 \title{\LARGE \bf
-Tibetan Machine Translation 
+Machine Translation for Tibetan Buddhist Texts
 }
@@
 
-\subsection{Preprocessing}
+\subsection{Data and Pre-processing}
 
-(Chris)
+To train our model, we acquired publicly available English to Tibetan parallel corpuses.
+The non-profit 84000 provided 80 English to Tibetan parallel texts on their website (citation).
+All 80 of these texts were expert translations of Indo-Tibtan Buddhist texts.
+For our model we used all 80 of these provided texts.
+\newline Before using them however, we pre-processed the provided data an a few ways.
+The provided files from 84000 were in a .tmx file format, a very standard parallel corpus file format.
+For our approach, we scraped all 80 of the .tmx file of both its English and Tibetan contents and dumped them into two large .txt files.
+One  .txt file contained the English lines, and the other .txt file contained the Tibetan translation.
+The total line numbers in the two .txt files were the same since one line in one file corresponding to exactly the same line of translated text in the other file.
 
@@
 (Jacob)
-
-\section{Data}
-
-(not sure if we need this section, maybe can be merged with another)
 
@@
 (Chris)
-(Rahul)
-(Kelsey)
 \begin{thebibliography}{99}

### 2021-03-27 paper/main.tex
added Ethics section

@@
 
+\section{ETHICAL CONSIDERATIONS}
+
+Care must be taken when approaching a culturally and spiritually important set of texts for use in a computational domain.
+To that end we authors acknowledge our place as outsiders of the Buddhist tradition and outline our ethical approach for translating these texts as outsiders.
+
+First, we acknowledge that it would be improper to consider our translations as genuine scripture.
+No matter what our results are, we authors are not in a qualified position to declare something as an honest translation of a Tibetan Buddhist text.
+To that end, what we have done is pass our results on to someone who can give a more qualified opinion of the accuracy of our translation an so we asked a Native Tibetan Speaker assess our results.
+This approach was inspired by Bird's Decolonizing NLP which suggests including a "human in the loop" is a key step in the process to decolonizing NLP research.
+
+Second, we acknowledge that there are certainly Indo-Tibetan texts that would be inappropriate to translate.
+These texts, the Tantric (esoteric) texts we understand and meant for a select subset of advanced practitioners and translating these texts for the masses could be spiritually dangerous.
+Knowing this, we have refrained from including and Tantric texts within our training data.
+
+Lastly, we acknowledge this is a tool to help bridge the language gap between practitioners both in and out of the Tibetan culture.
+We want to emphasize that this is just a tool.
+We make no claims that this tool will solve or fix the underlying social injustice issues within Tibetan community and diaspora.
+While we hope this tool can bring the community closer together and aid in prosperity, we make no claims that it will be able to do so.
+
 \section{PROCEDURE}
@@
 The non-profit 84000 provided 80 English to Tibetan parallel texts on their website (citation).
-All 80 of these texts were expert translations of Indo-Tibtan Buddhist texts.
+All 80 of these texts were expert translations of Indo-Tibetan Buddhist texts.
 For our model we used all 80 of these provided texts.
-\newline Before using them however, we pre-processed the provided data an a few ways.
+
+Before using them however, we pre-processed the provided data an a few ways.
 The provided files from 84000 were in a .tmx file format, a very standard parallel corpus file format.
@@
 
-Appendixes should appear before the acknowledgment. (not sure what we even need to put here)
+Appendixes should appear before the acknowledgment. (not sure what we even need to put here ~Chris)
 
@@
 
-(Thanks to professors + COMS 4995, etc etc)
-
-(Put in all papers from lit review here)
+The authors would like to thank Palden Gal for volunteering to read our translated texts.
+We would also like to thank Smara Muresan, Isabelle Zuagg, and Jonathan Reeve for mentoring this research.
+Lastly, we would like to thank the 84000 project for providing open source parallel Tibetan corpuses.
 

### 2021-03-31 paper/main.tex
Update main.tex
Added more to each of my sections

@@
 
-Care must be taken when approaching a culturally and spiritually important set of texts for use in a computational domain.
-To that end we authors acknowledge our place as outsiders of the Buddhist tradition and outline our ethical approach for translating these texts as outsiders.
+Care must be taken when approaching culturally and spiritually important texts for use in a computational domain.
+To that end we authors acknowledge our place as outsiders of the Buddhist tradition and outline our ethical approach for translating these texts as said outsiders.
 
@@
 No matter what our results are, we authors are not in a qualified position to declare something as an honest translation of a Tibetan Buddhist text.
-To that end, what we have done is pass our results on to someone who can give a more qualified opinion of the accuracy of our translation an so we asked a Native Tibetan Speaker assess our results.
-This approach was inspired by Bird's Decolonizing NLP which suggests including a "human in the loop" is a key step in the process to decolonizing NLP research.
+To that end, what we have done is pass our results on to someone who can give a more qualified opinion of the accuracy of our translation an so we asked a Native Tibetan Speaker who holds a Master's degree in Buddhist theology and philosophy to  assess our results.
+This approach was inspired by Bird's paper [citation:Decolonizing NLP ] which suggests including a "human in the loop" as a key step in the process to decolonizing NLP research.
 
 Second, we acknowledge that there are certainly Indo-Tibetan texts that would be inappropriate to translate.
-These texts, the Tantric (esoteric) texts we understand and meant for a select subset of advanced practitioners and translating these texts for the masses could be spiritually dangerous.
+These texts, namely the Tantric (esoteric) texts we understand are meant for a select subset of advanced practitioners and translating these texts for the masses could be spiritually dangerous.
 Knowing this, we have refrained from including and Tantric texts within our training data.
@@
 
-Before using them however, we pre-processed the provided data an a few ways.
-The provided files from 84000 were in a .tmx file format, a very standard parallel corpus file format.
-For our approach, we scraped all 80 of the .tmx file of both its English and Tibetan contents and dumped them into two large .txt files.
+Before using 84000's corpuses however, we pre-processed the provided data to meet our model's need.
+The provided files from 84000 were in a .tmx file format, a very standard parallel corpus file format [citation needed].
+For our approach, we scraped all 80 .tmx files for both its English and Tibetan contents and placed them line by line into two large .txt files.
 One .txt file contained the English lines, and the other .txt file contained the Tibetan translation.
-The total line numbers in the two .txt files were the same since one line in one file corresponding to exactly the same line of translated text in the other file.
+We made careful assurance that the lines remained in exactly one to one correspondance with their translations.
+That is, a newline charcter was used to distinguish two seperate lines of translation, but the same line number in each file corresponds to the exact translation provided by 84000.
+Therefore, the total line numbers in the two .txt files were necessarily the same since one line in one file corresponding to exactly the same line of translated text in the other file.
+Two such .txt file was made for each of the 80 .tmx files, a Tibetan and an English .txt file for each.
+After all 160 .txt files were generated, two single large .txt file was generated by concatenating all the Tibetan .txt files and all the English .txt files.
+These last two .txt files are what was fed into the model.
+
+TODO: Actually do this/remove following lines.
+It was not strictly necessary to separate the creation of the large .txt files into the smaller components.
+However, giving the limited data we had this method proved useful to see how much individual texts contributed to the model (ie by elminating some texts in some runs).
 

### 2021-04-01 paper/main.tex
Update main.tex
Local commit
Added prelim. Intro

@@
 
-(Kelsey)
+Our project aims to increase the accessibility of Tibetan religious texts by training a machine translation model specifically to translate religious Tibetan texts to English.
+We have trained our models in PyTorch and have leveraged the BART architecture in the HuggingFrance transformers library to make the training and architecture more straightforward.
+By creating thesee models, we hope to motivate scholars to study not only these texts but Tibetan as a whole.
+Currently, while there has been some interest in machine translation for Tibetan-Chinese and vice versa, the same interest has not been reflected in Tibetan-English machine translation.
+From a deep dive into the open-source Tibetan NLP material, we were unable to find any projects attempting to translate Tibetan-English nor could we find papers specifically discussing Tibetan-English machine translation.
+By having a readily available set of translated discourses could create an on-ramp for further language uses because increased access could lead to increased interest, especially in the English-speaking community.
 

### 2021-04-06 paper/main.tex
Update main.tex
Updated .tex with (most) citations plus Jacob's changes

@@
 \usepackage[T1]{fontenc}
+\usepackage{graphicx}
 
@@
 
-First, we acknowledge that it would be improper to consider our translations as genuine scripture.
+First, we acknowledge that it would be improper to consider our translations as genuine sutras.
 No matter what our results are, we authors are not in a qualified position to declare something as an honest translation of a Tibetan Buddhist text.
@@
 \subsection{Model Design}
+We used a 6 layer Transformer model that used a multi headed attention mechanism and a switchout transform.
+Each layer of the transformer model can be divided into an encoder and decoder layer.
+The encoding layer consists first of a self attention layer, an attention aspect which focuses on the encoded sentence, and then outputs to a feed forward layer.
+The multi part of the multi headed attention means that each head will encode the information in parallel and then the results will be concatenated and passed to the decoder.
+Figure one demonstrates this process.
+\includegraphics[scale=.4]{Transformer.png}
+\caption{\\Figure 1: Overview of a Transformer model}
+\\
+\par
+To prevent overfitting and to augment the data, we added in a switchout transform which will randomly switch out words in the source and target sentence.
+This also has the added advantage of augmenting the data even further which is especially important for our problem as religious tibetan has a relatively small corpus, around 30 thousand words.
+We also attempted to use masking, however we found it to be ineffective.
+Masking hides tokens from the model by replacing the tokens with a <mask> token.
+Tibetan tokens are not separated in the same way that latin languages separate their tokens.
+For example,
+\begin{verbatim}
+“བྱང་ཆུབ་མངོན་པར་རྫོགས་པར་འཚང་རྒྱ་སྟེ།” 
+\end{verbatim}
+a line in our test corpus translates to “Having the altruistic intention of excellent bodhicitta.” The tibetan word is only one token.
+Therefore, the mask would not help the model learn at all, as the entire sentence would be marked as <mask> and there would not be enough context for our model to discover the meaning.
+As a result, we did not use any form of masking in our model.
+Switchout does not create the same problem as their is always a token to take its place.
+\par
+After implementing our transformer model with switchout we tuned hyper parameters to maximize our Transformer model.
+We manually tuned the following hyperparameters: dropout, label smoothing, rnn size, the number of transformer layers and the number of heads of the attention model.
+The final configuration used can be found in the appendix.
+\par
+We used four different metrics to evaluate our results: accuracy, predicted score, perplexity and bilingual evaluation understudy (BLEU).
+The validation accuracy tells us how well we are learning our training data, the predicted score tells us the average log likelihood for each word, the perplexity tells us how well the model predicts the data and the bleu score evaluates the quality of the translation.
+Together, these metrics provide us with a decent picture of how well our model is performed.
+
+\section{Results}
+\begin{tabular}{ |c| c| c| c| }
+ Validation & Ave. Pred. Score & PPL & BLEU \\ 
+ 40.32\% & -1.69 & 5.41 & 23.01\% \\
+
+\end{tabular}\\
 
@@
 
-\section{Results}
+\section{Discussion}
 
-(TODO)
+(Kelsey)
 
@@
 
-(TODO)
+(Rahul)
 
@@
 
-\bibitem{c1} G.
-O.
-Young, ``Synthetic structure of industrial plastics (Book style with paper title and editor),'' 	in Plastics, 2nd ed.
-vol.
-3, J.
-Peters, Ed.
-New York: McGraw-Hill, 1964, pp.
-15--64.
-\bibitem{c2} W.-K. Chen, Linear Networks and Systems (Book style).	Belmont, CA: Wadsworth, 1993, pp. 123--135.
-\bibitem{c3} H. Poor, An Introduction to Signal Detection and Estimation.   New York: Springer-Verlag, 1985, ch. 4.
-\bibitem{c4} B. Smith, ``An approach to graphs of linear forms (Unpublished work style),'' unpublished.
-\bibitem{c5} E. H. Miller, ``A note on reflector arrays (Periodical styleÑAccepted for publication),'' IEEE Trans. Antennas Propagat., to be publised.
-\bibitem{c6} J. Wang, ``Fundamentals of erbium-doped fiber amplifiers arrays (Periodical styleÑSubmitted for publication),'' IEEE J. Quantum Electron., submitted for publication.
-\bibitem{c7} C. J. Kaufman, Rocky Mountain Research Lab., Boulder, CO, private communication, May 1995.
-\bibitem{c8} Y.
-Yorozu, M.
-Hirano, K.
-Oka, and Y.
-Tagawa, ``Electron spectroscopy studies on magneto-optical media and plastic substrate interfaces(Translation Journals style),'' IEEE Transl.
-J.
-Magn.Jpn., vol.
-2, Aug.
-1987, pp.
-740--741 [Dig.
-9th Annu.
-Conf.
-Magnetics Japan, 1982, p.
-301].
-\bibitem{c9} M. Young, The Techincal Writers Handbook.  Mill Valley, CA: University Science, 1989.
-\bibitem{c10} J. U. Duncombe, ``Infrared navigationÑPart I: An assessment of feasibility (Periodical style),'' IEEE Trans. Electron Devices, vol. ED-11, pp. 34--39, Jan. 1959.
-\bibitem{c11} S.
-Chen, B.
-Mulgrew, and P.
-M.
-Grant, ``A clustering technique for digital communications channel equalization using radial basis function networks,'' IEEE Trans.
-Neural Networks, vol.
-4, pp.
-570--578, July 1993.
-\bibitem{c12} R. W. Lucky, ``Automatic equalization for digital communication,'' Bell Syst. Tech. J., vol. 44, no. 4, pp. 547--588, Apr. 1965.
-\bibitem{c13} S.
-P.
-Bingulac, ``On the compatibility of adaptive controllers (Published Conference Proceedings style),'' in Proc.
-4th Annu.
-Allerton Conf.
-Circuits and Systems Theory, New York, 1994, pp.
-8--16.
-\bibitem{c14} G. R. Faulhaber, ``Design of service systems with priority reservation,'' in Conf. Rec. 1995 IEEE Int. Conf. Communications, pp. 3--8.
-\bibitem{c15} W. D. Doyle, ``Magnetization reversal in films with biaxial anisotropy,'' in 1987 Proc. INTERMAG Conf., pp. 2.2-1--2.2-6.
-\bibitem{c16} G.
-W.
-Juette and L.
-E.
-Zeffanella, ``Radio noise currents n short sections on bundle conductors (Presented Conference Paper style),'' presented at the IEEE Summer power Meeting, Dallas, TX, June 22--27, 1990, Paper 90 SM 690-0 PWRS.
-\bibitem{c17} J. G. Kreifeldt, ``An analysis of surface-detected EMG as an amplitude-modulated noise,'' presented at the 1989 Int. Conf. Medicine and Biological Engineering, Chicago, IL.
-\bibitem{c18} J. Williams, ``Narrow-band analyzer (Thesis or Dissertation style),'' Ph.D. dissertation, Dept. Elect. Eng., Harvard Univ., Cambridge, MA, 1993. 
-\bibitem{c19} N. Kawasaki, ``Parametric study of thermal and chemical nonequilibrium nozzle flow,'' M.S. thesis, Dept. Electron. Eng., Osaka Univ., Osaka, Japan, 1993.
-\bibitem{c20} J. P. Wilkinson, ``Nonlinear resonant circuit devices (Patent style),'' U.S. Patent 3 624 12, July 16, 1990.
+\bibitem{c1} Cai Z.
+(2015) A Method for the Chinese-Tibetan Machine Translation System’s Syntactic Analysis.
+In: Wong W.
+(eds) Proceedings of the 4th International Conference on Computer Engineering and Networks.
+Lecture Notes in Electrical Engineering, vol 355.
+Springer, Cham.
+
+\bibitem{c2} Liu, H., Zhao, W., Yu, X., & Wu, J. (2016). A Chinese to Tibetan Machine Translation System with Multiple Translating Strategies. Himalayan Linguistics, 15(1).
+
+\bibitem{c3} Hill, Nathan & Meelen, Marieke. (2017). Segmenting and POS tagging Classical Tibetan using a Memory-Based Tagger. Himalayan Linguistics. 16. 10.5070/H916234501.
+
+\bibitem{c4} Faggionato, Christian  and Meelen, Marieke.
+(2019).
+Developing the Old Tibetan Treebank.
+INCOMA Ltd.
+Proceedings of the International Conference on Recent Advances in Natural Language Processing (RANLP 2019) 304--312
+
+\bibitem{c5} Orna Almogi, Lena Dankin, Nachum Dershowitz, Lior Wolf.
+A Hackathon for Classical Tibetan.Journal of Data Mining and Digital Humanities, Episciences.org, 2019, Special Issue on Computer-Aided Processing of Intertextuality in Ancient Languages.
+hal-01371751v3
+
+\bibitem{c6} Orna Almogi, Lena Dankin, Nachum Dershowitz, Lior Wolf.
+A Hackathon for Classical Tibetan.Journal of Data Mining and Digital Humanities, Episciences.org, 2019, Special Issue on Computer-Aided Processing of Intertextuality in Ancient Languages.
+hal-01371751v3
 

### 2021-04-06 paper/main.tex
Update main.tex

@@
 No matter what our results are, we authors are not in a qualified position to declare something as an honest translation of a Tibetan Buddhist text.
-To that end, what we have done is pass our results on to someone who can give a more qualified opinion of the accuracy of our translation an so we asked a Native Tibetan Speaker who holds a Master's degree in Buddhist theology and philosophy to  assess our results.
-This approach was inspired by Bird's paper [citation:Decolonizing NLP ] which suggests including a "human in the loop" as a key step in the process to decolonizing NLP research.
+To that end, we have instead given our translations to someone who can give a more qualified opinion of the accuracy of our translation.
+In this case, we asked a Native Tibetan Speaker who holds a Master's degree in Buddhist theology and philosophy and a current PhD candidate in Tibetan History to  assess our results.
+This approach to invite someone directly in the Tibetan community to assess our results was inspired by Bird's paper \cite{c7} which suggests including a "human in the loop" as a key step in the process to decolonizing NLP research.
 
 Second, we acknowledge that there are certainly Indo-Tibetan texts that would be inappropriate to translate.
-These texts, namely the Tantric (esoteric) texts we understand are meant for a select subset of advanced practitioners and translating these texts for the masses could be spiritually dangerous.
+These texts, namely the Tantric (esoteric) texts, we understand are meant for a select subset of advanced practitioners and translating these texts for the masses could be spiritually dangerous.
 Knowing this, we have refrained from including and Tantric texts within our training data.
@@
 Lastly, we acknowledge this is a tool to help bridge the language gap between practitioners both in and out of the Tibetan culture.
-We want to emphasize that this is just a tool.
+We authors want to emphasize that this is just a tool.
 We make no claims that this tool will solve or fix the underlying social injustice issues within Tibetan community and diaspora.
@@
 To train our model, we acquired publicly available English to Tibetan parallel corpuses.
-The non-profit 84000 provided 80 English to Tibetan parallel texts on their website (citation).
+The non-profit 84000 provided 80 English to Tibetan parallel texts on their website \cite{c8}.
 All 80 of these texts were expert translations of Indo-Tibetan Buddhist texts.
@@
 Before using 84000's corpuses however, we pre-processed the provided data to meet our model's need.
-The provided files from 84000 were in a .tmx file format, a very standard parallel corpus file format [citation needed].
+The provided files from 84000 were in a .tmx file format, a very standard parallel corpus file format \cite{c9}.
 For our approach, we scraped all 80 .tmx files for both its English and Tibetan contents and placed them line by line into two large .txt files.
 One .txt file contained the English lines, and the other .txt file contained the Tibetan translation.
-We made careful assurance that the lines remained in exactly one to one correspondance with their translations.
-That is, a newline charcter was used to distinguish two seperate lines of translation, but the same line number in each file corresponds to the exact translation provided by 84000.
+We made careful assurance that the lines remained in exactly one to one correspondence with their translations.
+That is, a newline character was used to distinguish two separate lines of translation, but the same line number in each file corresponds to the exact translation provided by 84000.
 Therefore, the total line numbers in the two .txt files were necessarily the same since one line in one file corresponding to exactly the same line of translated text in the other file.
-Two such .txt file was made for each of the 80 .tmx files, a Tibetan and an English .txt file for each.
+Two such .txt file were made for each of the 80 .tmx files, a Tibetan and an English .txt file for each.
 After all 160 .txt files were generated, two single large .txt file was generated by concatenating all the Tibetan .txt files and all the English .txt files.
@@
 It was not strictly necessary to separate the creation of the large .txt files into the smaller components.
-However, giving the limited data we had this method proved useful to see how much individual texts contributed to the model (ie by elminating some texts in some runs).
+However, giving the limited data we had this method proved useful to see how much individual texts contributed to the model (ie by eliminating some texts in some runs).
 
@@
 
+\bibitem{c7} Bird, S. (2020). Decolonising Speech and Language Technology. Proceedings of the 28th International Conference on Computational Linguistics (COLING 2020), 3504–3519.
+
+\bibitem{c8} 84000 Project (2021). What We Do. https://84000.co/about/what-we-do
+
+\bibitem{c9} Coverpages (1997). TMX Format. http://xml.coverpages.org/tmxSpec971212.html
+
 \end{thebibliography}

### 2021-04-07 paper/main.tex
Edited Intro and added to Disc.

@@
 
+Digital language justice is the right for an user to use the language they are most comfortable with in the digital sphere.
+Language justice goes beyond this, however.
+With modern technology, it is possible to reconnect people with their cultural dispora.
 Our project aims to increase the accessibility of Tibetan religious texts by training a machine translation model specifically to translate religious Tibetan texts to English.
-We have trained our models in PyTorch and have leveraged the BART architecture in the HuggingFrance transformers library to make the training and architecture more straightforward.
+We hope to create a translation device for Tibetan Buddhist practitioners who are non-Tibetan speakers, allowing them access to texts that had previously been barred by the language barrier.
+We have trained our model in PyTorch and have leveraged the BART architecture in the HuggingFrance transformers library to make the training and architecture more straightforward.
 By creating thesee models, we hope to motivate scholars to study not only these texts but Tibetan as a whole.
@@
 
+Considering our use of word segmentation for preprocessing our data (as opposed to character segmentation), our results are comparable with, albeit lower than, Tibetan-Chinese models.
+Lai, Zhao, and Bao ran several different Tibetan-Chinese models on the same data using two different preprocessing methodologies, character segmentation and word segmentation.
+Their word segmented Transformer model averaged a higher BLEU score than ours, scoring a 38.44\% while our model scored 23.01\%.
+
 (Kelsey)

### 2021-04-07 paper/main.tex
added results section

@@
 
-(Jacob)
+\par
+Our six layer transformer model with switch out was unsuccessful at translating religious tibetan to english.
+The translated output does not match closely enough in either pure accuracy or in the meaning of the passage to be useful in translating religious tibetan texts to english.
+\par
+While superficially, the perplexity appears to demonstrate a strong model, further investigation demonstrates that the low perplexity is a sign of inaccuracy.
+Perplexity measures the number of possible words that a model could predict for any given sentence.
+A higher perplexity indicates higher uncertainty.
+However, the low BLEU score, which measures the quality of the machine translation, has already demonstrated that the accuracy is low.
+This combined with the perplexity shows the models overconfidence in predicting certain phrases.
+For example, the tibetan phrase །དེ་ནས་བཅོམ་ལྡན་འདས་ལ་བྱང་ཆུབ་སེམས་དཔའ་སེམས་དཔའ་ཆེན་པོ་བྱམས་པས་འདི་སྐད་ཅེས་གསོལ་ཏོ། translates to “So the bodhisattva mahāsattva Maitreya asked the Blessed One”, but the top five of the predicted translations all are of the form “The Blessed One then said to the [adj] [proper noun]”.
+Even though this sentence is wrong, the perplexity is low because the model is confident in the first part of the sentence.
+This phenomenon is repeated throughout the results.
+Keyphrases from the Tibetan sentence are interpreted wrongly in the same way, and then the model learns to be confident about its predictions which lowers the perplexity.
+\par
+The results demonstrate the flaw in the Transformer model using the current parallel Tibetan to English corpuses.
+Each token in the Tibetan sentence corresponds to an entire phrase in english rather than a separate word.
+This complicates the translation, and the model learns parts of the phrases quite well.
+The validation accuracy is relatively high because the transformer model is good at learning parts of these phrases but it is not good at learning entire sentences or stringing together sentences in context which contributes to the low BLEU score.
+Overall, the model is capable of learning specific phrases well, but fails at the broad task of creating coherent translations from religious tibetan texts to english.
 

### 2021-04-13 paper/main.tex
First draft of discussion

@@
 Their word segmented Transformer model averaged a higher BLEU score than ours, scoring a 38.44\% while our model scored 23.01\%.
+While this comparisons may lead to the conclusion that our model is flawed and irredeemable, only the former is correct.
+There’s no doubt that our model is underperforming in comparison to other transformer models and hardly makes an correct predictions at all.
+However, this is most likely due to our difference in preprocessing techniques and the linguistic differences and similarities between English, Chinese, and Tibetan.
+
+Unlike English but similar to Chinese, Tibetan words are not clearly separated by spaces, rather, characters which represent words are strung together to make a sentence.
+This makes preprocessing the data and mapping Tibetan words to their English counterparts a difficult task and it makes preprocessing on characters impossible.
+We trained our model on a parallel corpus of English and Tibetan sentences instead of segmenting each word since preprocessing these sentences to word-to-word mappings would require an extensive knowledge on the Tibetan language which our team did not have.
+On top of this, a Chinese-Tibetan model might perform overall better than a Tibetan-English one since Chinese and Tibetan are part of the same language family, and it is likely that characters may map to one another and the same preprocessing techniques can be used on both languages.
+Therefore, while comparing our model to Lai, Zhao and Bao’s model may be helpful to give a general baseline on how other Tibetan translators are performing, it should not be considered where our model must perform.
+
+Another issue with our data that may be remedied in future iterations is about the corpora we chose to train and test our model on.
+Since we trained on sentences in our testing set, it is very possible that the model may have read some of the initial characters and assumed that the entire sentence would be a sentence from our training data that began with the same characters.
+This may explain why our model overpredicts the sentence “Our blessed one” with such a low perplexity.
+As stated above, this issue might be remedied if we could preprocess our data so that words mapped to words but since we our limited on our ability to do such, we could consider training our model on a different corpus altogether.
+Instead of using the translated Buddhist Sutras to train our model, we could use a more general Tibetan-English corpus to train our model and then use transfer learning to finetune our model to religious texts.
+However, at this current time, the translated sutras provided by 84000 are some of the only parallel translations available on the internet.
 

### 2021-04-13 paper/main.tex
Update main.tex

@@
 
-\documentclass[letterpaper, 10 pt, conference]{ieeeconf}  % Comment this line out
+\documentclass[letterpaper, 12 pt, conference]{ieeeconf}  % Comment this line out
                                                           % if you need a4paper
@@
 \author{ Jacob Atkins$^{1}$, Chris Calloway$^{2}$, Kelsey Namba$^{3}$, Rahul Sehrawat$^{4}$% <-this % stops a space
-\thanks{*Smara Muresan, Isabelle Zuagg, Jonathan Reeve}% <-this % stops a space
+\thanks{* Supervised by Smara Muresan, Isabelle Zuagg, and Jonathan Reeve}% <-this % stops a space
+\thanks{$^{1}$ Jacob Atkins, Columbia '21
+        {\tt\small jha2137 at columbia dot edu}}%
+
+\thanks{$^{2}$ Chris Calloway, Columbia '21
+        {\tt\small cmc2374 at columbia dot edu}}%
+
+\thanks{$^{3}$ Kelsey Namba, Columbia '21
+        {\tt\small kn2426 at columbia dot edu}}%
+
+\thanks{$^{3}$ Rahul Sehrawat Columbia '21
+        {\tt\small rs3688 at columbia dot edu}}%
 
@@
 
-(Rahul)
+(Rahul) "PLACEHOLDER PLACEHOLDER PLACEHOLDER PLACEHOLDER PLACEHOLDER PLACEHOLDER PLACEHOLDER PLACEHOLDER PLACEHOLDER PLACEHOLDER PLACEHOLDER
+PLACEHOLDER PLACEHOLDER PLACEHOLDER PLACEHOLDER PLACEHOLDER
+PLACEHOLDER PLACEHOLDER PLACEHOLDER PLACEHOLDER PLACEHOLDER
+PLACEHOLDER PLACEHOLDER PLACEHOLDER PLACEHOLDER PLACEHOLDER
+PLACEHOLDER PLACEHOLDER PLACEHOLDER PLACEHOLDER PLACEHOLDER
+PLACEHOLDER PLACEHOLDER PLACEHOLDER PLACEHOLDER PLACEHOLDER
 
@@
 
-Digital language justice is the right for an user to use the language they are most comfortable with in the digital sphere.
-Language justice goes beyond this, however.
-With modern technology, it is possible to reconnect people with their cultural dispora.
 Our project aims to increase the accessibility of Tibetan religious texts by training a machine translation model specifically to translate religious Tibetan texts to English.
-We hope to create a translation device for Tibetan Buddhist practitioners who are non-Tibetan speakers, allowing them access to texts that had previously been barred by the language barrier.
-We have trained our model in PyTorch and have leveraged the BART architecture in the HuggingFrance transformers library to make the training and architecture more straightforward.
-By creating thesee models, we hope to motivate scholars to study not only these texts but Tibetan as a whole.
+We have trained our models in PyTorch and have leveraged the BART architecture in the HuggingFrance transformers library to make the training and architecture more straightforward.
+By creating these models, we hope to motivate scholars to study not only these texts but Tibetan as a whole.
 Currently, while there has been some interest in machine translation for Tibetan-Chinese and vice versa, the same interest has not been reflected in Tibetan-English machine translation.
@@
 By having a readily available set of translated discourses could create an on-ramp for further language uses because increased access could lead to increased interest, especially in the English-speaking community.
+
+\subsection{RELATED WORK}
+
+As a whole, not much work has been done in the area of Machine Translation or even Natural Language Processing as a whole for the Tibetan Language.
+There are a few papers exist which concern morphological segmentation and POS tagging for the Tibetan language \cite{c3} \cite{c4}.
+In recent years there has considerable interest in Chinese to Tibetan Machine translation \cite{c1} \cite{c2}.
+For example, “A Chinese to Tibetan Machine Translation System with Multiple Translating Strategies” discusses an approach of machine translation from Tibetan to Chinese.
+The paper makes 3 different translations models, Ruled Based MT, Example Based MT, and Statistical Based MT,  and compares the results.
+Noticeably missing from this is a Neural Network MT.
+However, Neural Network MT wasn’t prevalent at that time of its authorship.
+Nonetheless, the paper concludes, given this data (after processing), that the Statistical Based MT (implemented with conditional random fields) is indeed the best performing option.
+This is perhaps to be expected.
+However, as is alluded to above it is unknown whether a Neural Network MT would give a better performance here.
+
+The paper “Tibetan-Chinese Neural Machine Translation based on Syllable Segmentation” explores four different methodologies for Tibetan to Chinese machine translation.
+Chinese and Tibetan are genetically related sharing the same proto-family, the Sino-Tibetan language family.
+This may mean that the results of this paper may not be fully reflected within our own project.
+However, there are several important findings that Lai, Zhao, and Bao discovered over the course of their experiment, particularly on the impacts of different models.
+In 2018, the statistical models were the preferred model type for Tibetan translation tasks, based on the linguistic characteristics of Tibetan.
+Lai, Zhao, and Bao test whether this is truly the most accurate and efficient model for these tasks by testing Seq2Seq, RNNSearch, Fairseq, and Transformer models against one another.
+Although each model has its own specific architecture that makes it difficult to normalize settings across all four, the researchers had the same basic parameters (same vocabulary size and training iterations).
+They also trained each model on two different corpora, one processed based on word segmentation and one processed based on character segmentation (or syllable segmentation).
+Each model’s performance scores were measured by the BLEU index.
+
+The first model discussed is the Seq2Seq model, or a sequence to sequence model, which is a model that takes in a sequence as input and outputs a sequence as well.
+The basic idea of a Seq2Seq model is to map the input sequence to the output sequence by taking the input through an LSTM which creates a vector of weighted sums during  the encoding process and returns the words with the highest probability of occurring in the decoder process.
+The second model tested is the RNNSearch model that adds an attention mechanism, a mechanism that retains the values calculated by the encoding process to train another model which associates the values with output sequences.
+The new model produces word vectors and then uses these vectors to produce the next word.
+Although Fairseq models essentially have the same structure as the RNNSearch, where both employ encoder-decoders with attention mechanisms, Fairseq models use CNNs to map the input to the output sequence, which also allows for all the calculations of input sequences to run in parallel.
+The final model that they tested was the transformer model, which uses the attention mechanism without a CNN or RNN structure.
+Instead, transformers use a self-attention mechanism that relates all the words in the sentence to one another, where order doesn't matter.
+Out of these four models, the transformer model trained on the syllable parsed corpus performed the best.
+For each model tested, they performed substantially better with the syllable parsed corpus than the word parsed corpus.
+This means that one way we might look to improve our own project’s accuracy is to use a character segmented corpus rather than a word segmented corpus.
+
+No such papers exist for Tibetan to Enlgish Machine transltions.
+In fact, to our knowledge this is the first such Tibetan to English Machine translation system developed.
+Nonetheless, the findings from the Chinese to Tibetan machine translation have given us a starting point for our investitagion.
 
@@
 First, we acknowledge that it would be improper to consider our translations as genuine sutras.
-No matter what our results are, we authors are not in a qualified position to declare something as an honest translation of a Tibetan Buddhist text.
+Regardless of our quantitative results, we authors are not in a qualified position to declare something as an honest translation of a Tibetan Buddhist text.
 To that end, we have instead given our translations to someone who can give a more qualified opinion of the accuracy of our translation.
-In this case, we asked a Native Tibetan Speaker who holds a Master's degree in Buddhist theology and philosophy and a current PhD candidate in Tibetan History to  assess our results.
-This approach to invite someone directly in the Tibetan community to assess our results was inspired by Bird's paper \cite{c7} which suggests including a "human in the loop" as a key step in the process to decolonizing NLP research.
+To this end, we asked a Native Tibetan Speaker who holds a Master's degree in Buddhist theology and philosophy and a current PhD candidate in Tibetan History to  assess our results.
+This approach to invite someone directly in the Tibetan community to assess our results was inspired by Bird's paper \cite{c7} which suggests including a "human in the loop" as a key step in the process to decolonizing Natural Language Processing research.
 
@@
 
-To train our model, we acquired publicly available English to Tibetan parallel corpuses.
+To train our model, we acquired publicly available English to Tibetan parallel corpora.
 The non-profit 84000 provided 80 English to Tibetan parallel texts on their website \cite{c8}.
@@
 
-Before using 84000's corpuses however, we pre-processed the provided data to meet our model's need.
+Before using 84000's corpora however, we pre-processed the provided data to meet our model's need.
 The provided files from 84000 were in a .tmx file format, a very standard parallel corpus file format \cite{c9}.
@@
 
-TODO: Actually do this/remove following lines.
 It was not strictly necessary to separate the creation of the large .txt files into the smaller components.
-However, giving the limited data we had this method proved useful to see how much individual texts contributed to the model (ie by eliminating some texts in some runs).
+However, giving the limited data we had this method proved useful to see how much individual texts contributed to the model.
 
@@
 To prevent overfitting and to augment the data, we added in a switchout transform which will randomly switch out words in the source and target sentence.
-This also has the added advantage of augmenting the data even further which is especially important for our problem as religious tibetan has a relatively small corpus, around 30 thousand words.
+This also has the added advantage of augmenting the data even further which is especially important for our problem as religious Tibetan has a relatively small corpus, around 30 thousand words.
 We also attempted to use masking, however we found it to be ineffective.
@@
 \end{verbatim}
-a line in our test corpus translates to “Having the altruistic intention of excellent bodhicitta.” The tibetan word is only one token.
+a line in our test corpus translates to “Having the altruistic intention of excellent bodhicitta.” The Tibetan word is only one token.
 Therefore, the mask would not help the model learn at all, as the entire sentence would be marked as <mask> and there would not be enough context for our model to discover the meaning.
@@
 
-\par
-Our six layer transformer model with switch out was unsuccessful at translating religious tibetan to english.
-The translated output does not match closely enough in either pure accuracy or in the meaning of the passage to be useful in translating religious tibetan texts to english.
-\par
-While superficially, the perplexity appears to demonstrate a strong model, further investigation demonstrates that the low perplexity is a sign of inaccuracy.
-Perplexity measures the number of possible words that a model could predict for any given sentence.
-A higher perplexity indicates higher uncertainty.
-However, the low BLEU score, which measures the quality of the machine translation, has already demonstrated that the accuracy is low.
-This combined with the perplexity shows the models overconfidence in predicting certain phrases.
-For example, the tibetan phrase །དེ་ནས་བཅོམ་ལྡན་འདས་ལ་བྱང་ཆུབ་སེམས་དཔའ་སེམས་དཔའ་ཆེན་པོ་བྱམས་པས་འདི་སྐད་ཅེས་གསོལ་ཏོ། translates to “So the bodhisattva mahāsattva Maitreya asked the Blessed One”, but the top five of the predicted translations all are of the form “The Blessed One then said to the [adj] [proper noun]”.
-Even though this sentence is wrong, the perplexity is low because the model is confident in the first part of the sentence.
-This phenomenon is repeated throughout the results.
-Keyphrases from the Tibetan sentence are interpreted wrongly in the same way, and then the model learns to be confident about its predictions which lowers the perplexity.
-\par
-The results demonstrate the flaw in the Transformer model using the current parallel Tibetan to English corpuses.
-Each token in the Tibetan sentence corresponds to an entire phrase in english rather than a separate word.
-This complicates the translation, and the model learns parts of the phrases quite well.
-The validation accuracy is relatively high because the transformer model is good at learning parts of these phrases but it is not good at learning entire sentences or stringing together sentences in context which contributes to the low BLEU score.
-Overall, the model is capable of learning specific phrases well, but fails at the broad task of creating coherent translations from religious tibetan texts to english.
+(Jacob)
 
@@
 
-\section*{APPENDIX}
-
-Appendixes should appear before the acknowledgment. (not sure what we even need to put here ~Chris)
-
 \section*{ACKNOWLEDGMENT}
 
-The authors would like to thank Palden Gal for volunteering to read our translated texts.
+The authors would like to thank Palden Gyal for volunteering to read our translated texts.
 We would also like to thank Smara Muresan, Isabelle Zuagg, and Jonathan Reeve for mentoring this research.
-Lastly, we would like to thank the 84000 project for providing open source parallel Tibetan corpuses.
+Lastly, we would like to thank the 84000 project for providing open source parallel Tibetan corpora.
 

### 2021-04-13 paper/main.tex
Fixed grammar errors , still need to cite

@@
 Lai, Zhao, and Bao ran several different Tibetan-Chinese models on the same data using two different preprocessing methodologies, character segmentation and word segmentation.
-Their word segmented Transformer model averaged a higher BLEU score than ours, scoring a 38.44\% while our model scored 23.01\%.
-While this comparisons may lead to the conclusion that our model is flawed and irredeemable, only the former is correct.
-There’s no doubt that our model is underperforming in comparison to other transformer models and hardly makes an correct predictions at all.
+Their word segmented Transformer model averaged a higher BLEU score than ours, scoring 38.44\% while our model scored 23.01\%.
+While this comparison may lead to the conclusion that our model is flawed and irredeemable, only the former is correct.
+There’s no doubt that our model is underperforming in comparison to other transformer models and hardly makes a correct prediction at all.
 However, this is most likely due to our difference in preprocessing techniques and the linguistic differences and similarities between English, Chinese, and Tibetan.
 
-Unlike English but similar to Chinese, Tibetan words are not clearly separated by spaces, rather, characters which represent words are strung together to make a sentence.
+Unlike English but similar to Chinese, Tibetan words are not clearly separated by spaces, rather, characters that represent words are strung together to make a sentence.
 This makes preprocessing the data and mapping Tibetan words to their English counterparts a difficult task and it makes preprocessing on characters impossible.
-We trained our model on a parallel corpus of English and Tibetan sentences instead of segmenting each word since preprocessing these sentences to word-to-word mappings would require an extensive knowledge on the Tibetan language which our team did not have.
-On top of this, a Chinese-Tibetan model might perform overall better than a Tibetan-English one since Chinese and Tibetan are part of the same language family, and it is likely that characters may map to one another and the same preprocessing techniques can be used on both languages.
-Therefore, while comparing our model to Lai, Zhao and Bao’s model may be helpful to give a general baseline on how other Tibetan translators are performing, it should not be considered where our model must perform.
+We trained our model on a parallel corpus of English and Tibetan sentences instead of segmenting each word since preprocessing these sentences to word-to-word mappings would require extensive knowledge of the Tibetan language which our team did not have.
+On top of this, a Chinese-Tibetan model might perform overall better than a Tibetan-English one since Chinese and Tibetan are part of the same language family, and it is likely that characters may map to one another and the same preprocessing techniques can be used in both languages.
+Therefore, while comparing our model to Lai, Zhao, and Bao’s model may be helpful to give a general baseline on how other Tibetan translators are performing, it should not be considered where our model must perform.
 
-Another issue with our data that may be remedied in future iterations is about the corpora we chose to train and test our model on.
+Another issue with our data that may be remedied in future iterations is the corpora we chose to train and test our model on.
 Since we trained on sentences in our testing set, it is very possible that the model may have read some of the initial characters and assumed that the entire sentence would be a sentence from our training data that began with the same characters.
 This may explain why our model overpredicts the sentence “Our blessed one” with such a low perplexity.
-As stated above, this issue might be remedied if we could preprocess our data so that words mapped to words but since we our limited on our ability to do such, we could consider training our model on a different corpus altogether.
-Instead of using the translated Buddhist Sutras to train our model, we could use a more general Tibetan-English corpus to train our model and then use transfer learning to finetune our model to religious texts.
+As stated above, this issue might be remedied if we could preprocess our data so that words mapped to words but since we are limited by our ability to do such, we could consider training our model on a different corpus altogether.
+Instead of using the translated Buddhist sutras to train our model, we could use a more general Tibetan-English corpus to train our model and then use transfer learning to fine-tune our model to religious texts.
 However, at this current time, the translated sutras provided by 84000 are some of the only parallel translations available on the internet.
+
+Perhaps the best way to translate Tibetan to English would be to take another approach to our model design altogether and abandon the tested Transformer model.
+We could build an Encoder-Decoder structure based on Back Translation which would help deal with the lack of Tibetan-English translations.
+In this model, we would mix Tibetan and English sentences together to train the model.
+This would help to increase the size of our training data since it would no longer just be Tibetan sentences.
+Since this type of model has yet to be tested for Tibetan-English, it's uncertain whether this change could actually improve our translation accuracy and BLEU score, however based on papers written on Back Translation, this design looks promising.
 

### 2021-04-14 paper/main.tex
Update main.tex

@@
 
-(Rahul) "PLACEHOLDER PLACEHOLDER PLACEHOLDER PLACEHOLDER PLACEHOLDER PLACEHOLDER PLACEHOLDER PLACEHOLDER PLACEHOLDER PLACEHOLDER PLACEHOLDER
-PLACEHOLDER PLACEHOLDER PLACEHOLDER PLACEHOLDER PLACEHOLDER
-PLACEHOLDER PLACEHOLDER PLACEHOLDER PLACEHOLDER PLACEHOLDER
-PLACEHOLDER PLACEHOLDER PLACEHOLDER PLACEHOLDER PLACEHOLDER
-PLACEHOLDER PLACEHOLDER PLACEHOLDER PLACEHOLDER PLACEHOLDER
-PLACEHOLDER PLACEHOLDER PLACEHOLDER PLACEHOLDER PLACEHOLDER
+This paper’s presents our attempt at training a model focusing on the translation of religious Tibetan texts to the English language.
+Tibetan is an extremely low-resource language and presents its own unique challenges in every stage toward the development of a high accuracy model.
+We look to explore ethical considerations of translating religious texts, provide an in-depth view of our procedure and follow up with discussions, results and conclusions.
+In support, we utilize various resources including conversations with a native speaker and the exploration of open source parallel Tibetan corpuses.
 
@@
 
-(Rahul)
+As mentioned in the beginning, developing a machine learning model for Tibetan Buddhist texts is a challenging case study primarily due to the extremely low resources available.
+We overcame many obstacles while training our model and even employed the usefulness of native speakers along with publicly available resources, but our 6 layer transformer model with switch out was unsuccessful at translating religious Tibetan to English.
+The translated output does not match closely enough in either pure accuracy or in the meaning of the passage to be useful in translating religious Tibetan texts to English, as expanded upon in the results section.
+Additionally, we explored the ethical considerations of translating such texts and deeply reflected on those, amongst our procedure and findings.
 

### 2021-04-15 paper/main.tex
Added Citation

@@
 Considering our use of word segmentation for preprocessing our data (as opposed to character segmentation), our results are comparable with, albeit lower than, Tibetan-Chinese models.
-Lai, Zhao, and Bao ran several different Tibetan-Chinese models on the same data using two different preprocessing methodologies, character segmentation and word segmentation.
+Lai, Zhao, and Bao \cite{2} ran several different Tibetan-Chinese models on the same data using two different preprocessing methodologies, character segmentation and word segmentation.
 Their word segmented Transformer model averaged a higher BLEU score than ours, scoring 38.44\% while our model scored 23.01\%.
@@
 
-Unlike English but similar to Chinese, Tibetan words are not clearly separated by spaces, rather, characters that represent words are strung together to make a sentence.
+Unlike English but similar to Chinese, Tibetan words are not clearly separated by spaces, rather, characters that represent words are strung together to make a sentence \cite{5}.
 This makes preprocessing the data and mapping Tibetan words to their English counterparts a difficult task and it makes preprocessing on characters impossible.
@@
 On top of this, a Chinese-Tibetan model might perform overall better than a Tibetan-English one since Chinese and Tibetan are part of the same language family, and it is likely that characters may map to one another and the same preprocessing techniques can be used in both languages.
-Therefore, while comparing our model to Lai, Zhao, and Bao’s model may be helpful to give a general baseline on how other Tibetan translators are performing, it should not be considered where our model must perform.
+Therefore, while comparing our model to Lai, Zhao, and Bao’s \cite{2} model may be helpful to give a general baseline on how other Tibetan translators are performing, it should not be considered where our model must perform.
 
@@
 Instead of using the translated Buddhist sutras to train our model, we could use a more general Tibetan-English corpus to train our model and then use transfer learning to fine-tune our model to religious texts.
-However, at this current time, the translated sutras provided by 84000 are some of the only parallel translations available on the internet.
+However, at this current time, the translated sutras provided by 84000 \cite{8} are some of the only parallel translations available on the internet.
 
@@
 We could build an Encoder-Decoder structure based on Back Translation which would help deal with the lack of Tibetan-English translations.
-In this model, we would mix Tibetan and English sentences together to train the model.
+In this model, we would mix Tibetan and English sentences together to train the model \cite{10}.
 This would help to increase the size of our training data since it would no longer just be Tibetan sentences.
 Since this type of model has yet to be tested for Tibetan-English, it's uncertain whether this change could actually improve our translation accuracy and BLEU score, however based on papers written on Back Translation, this design looks promising.
-
-(Kelsey)
 
@@
 
+\bibitem{c10} Ma, Edward. (2019). Data Augmentation in NLP. https://towardsdatascience.com/data-augmentation-in-nlp-2801a34dfc28
+
 \end{thebibliography}

### 2021-04-15 paper/main.tex
Added more to the introduction

@@
 By creating these models, we hope to motivate scholars to study not only these texts but Tibetan as a whole.
+Having a readily available set of translated discourses could create an on-ramp for further language uses because increased access could lead to increased interest, especially in the English-speaking community.
 Currently, while there has been some interest in machine translation for Tibetan-Chinese and vice versa, the same interest has not been reflected in Tibetan-English machine translation.
 From a deep dive into the open-source Tibetan NLP material, we were unable to find any projects attempting to translate Tibetan-English nor could we find papers specifically discussing Tibetan-English machine translation.
-By having a readily available set of translated discourses could create an on-ramp for further language uses because increased access could lead to increased interest, especially in the English-speaking community.
+While our model tackles this difficult task in creating a translator for Tibetan-English, it is by no means a perfect or even accurate model.
+In the final iteration of our model, we were only able to achieve a 23.01% BLEU score, which means that our model has low-quality output.
+This is caused by several factors, from the difficulty to preprocess Tibetan sentences to the lack of Tibetan-English data freely available online and our overall knowledge of the Tibetan language.
+While it may seem that we were unable to fully accomplish our lofty goal of creating a Tibetan-English translator it’s important to note that we are treading in nearly uncharted territory.
+We hope that our endeavor may help future projects in this field and provide insight into the current state of Tibetan-English machine translation.
 

### 2021-04-15 paper/main.tex
removed tibetan text

@@
 For example,
-\begin{verbatim}
-“བྱང་ཆུབ་མངོན་པར་རྫོགས་པར་འཚང་རྒྱ་སྟེ།” 
-\end{verbatim}
-a line in our test corpus translates to “Having the altruistic intention of excellent bodhicitta.” The Tibetan word is only one token.
+a line in our test corpus translated to “Having the altruistic intention of excellent bodhicitta.” However, the Tibetan word is only one token.
 Therefore, the mask would not help the model learn at all, as the entire sentence would be marked as <mask> and there would not be enough context for our model to discover the meaning.
@@
 This combined with the perplexity shows the models overconfidence in predicting certain phrases.
-For example, the tibetan phrase །དེ་ནས་བཅོམ་ལྡན་འདས་ལ་བྱང་ཆུབ་སེམས་དཔའ་སེམས་དཔའ་ཆེན་པོ་བྱམས་པས་འདི་སྐད་ཅེས་གསོལ་ཏོ། translates to “So the bodhisattva mahāsattva Maitreya asked the Blessed One”, but the top five of the predicted translations all are of the form “The Blessed One then said to the [adj] [proper noun]”.
+For example, one Tibetan phrases expert translation was “So the bodhisattva mahāsattva Maitreya asked the Blessed One”, but the top five of the predicted translations all are of the form “The Blessed One then said to the [adj] [proper noun]”.
 Even though this sentence is wrong, the perplexity is low because the model is confident in the first part of the sentence.

### 2021-04-16 paper/main.tex
cited image
Update main.tex
Fixed typos in abstract

@@
 
-This paper’s presents our attempt at training a model focusing on the translation of religious Tibetan texts to the English language.
-Tibetan is an extremely low-resource language and presents its own unique challenges in every stage toward the development of a high accuracy model.
-We look to explore ethical considerations of translating religious texts, provide an in-depth view of our procedure and follow up with discussions, results and conclusions.
-In support, we utilize various resources including conversations with a native speaker and the exploration of open source parallel Tibetan corpuses.
+This paper presents our attempt at training a model focusing on the translation of religious Tibetan texts to the English language.
+Tibetan is an extremely low-resource language and presents its own unique challenges in all stages of the development of a high accuracy machine translation model.
+In this paper, we explore the ethical considerations of translating religious texts, provide an in-depth view of our model creation procedure and follow up with discussions, results and conclusions.
+In support, we utilize various resources including scheduling conversations with a native speaker and the exploration of open source parallel Tibetan corpora.
 
@@
 \includegraphics[scale=.4]{Transformer.png}
-\caption{\\Figure 1: Overview of a Transformer model}
+\caption{\\Figure 1: Overview of a Transformer model \cite{11}}
 \\
@@
 
+\bibitem{c11} Doshi, Ketan (2021).
+Transformers Explained Visually (Part 3): Multi-head Attention, deep dive.
+https://towardsdatascience.com/transformers-explained-visually-part-3-multi-head-attention-deep-dive-1c1ff1024853
+
 \end{thebibliography}

### 2021-04-16 paper/main.tex
Update main.tex
Added a sentence or two to the into. Fixed a bug with a "%" that now makes us 6 pages

@@
 Our project aims to increase the accessibility of Tibetan religious texts by training a machine translation model specifically to translate religious Tibetan texts to English.
+Tibetan religions, in particular Buddhism, have a long tradition of translation of texts.
+While this work was not created within the context of theses traditions, it was certainly inspired by them and we hope it inspires more such translations.
+
 We have trained our models in PyTorch and have leveraged the BART architecture in the HuggingFrance transformers library to make the training and architecture more straightforward.
@@
 Currently, while there has been some interest in machine translation for Tibetan-Chinese and vice versa, the same interest has not been reflected in Tibetan-English machine translation.
-From a deep dive into the open-source Tibetan NLP material, we were unable to find any projects attempting to translate Tibetan-English nor could we find papers specifically discussing Tibetan-English machine translation.
+We attempted to find open-source Tibetan NLP material, but we were unable to find any projects attempting to translate Tibetan-English nor could we find papers specifically discussing Tibetan-English machine translation.
 While our model tackles this difficult task in creating a translator for Tibetan-English, it is by no means a perfect or even accurate model.
-In the final iteration of our model, we were only able to achieve a 23.01% BLEU score, which means that our model has low-quality output.
+In the final iteration of our model, we were only able to achieve a 23.01\% BLEU score, which means that our model has low-quality output.
+
 This is caused by several factors, from the difficulty to preprocess Tibetan sentences to the lack of Tibetan-English data freely available online and our overall knowledge of the Tibetan language.

### 2021-04-16 paper/main.tex
Update main.tex
Fixed citation typos, removed appendix sentence, and other small changes

@@
 
-\thanks{$^{3}$ Rahul Sehrawat Columbia '21
+\thanks{$^{4}$ Rahul Sehrawat Columbia '21
         {\tt\small rs3688 at columbia dot edu}}%
@@
 
+\subsection{Overview}
+
 Our project aims to increase the accessibility of Tibetan religious texts by training a machine translation model specifically to translate religious Tibetan texts to English.
@@
 We attempted to find open-source Tibetan NLP material, but we were unable to find any projects attempting to translate Tibetan-English nor could we find papers specifically discussing Tibetan-English machine translation.
+
 While our model tackles this difficult task in creating a translator for Tibetan-English, it is by no means a perfect or even accurate model.
 In the final iteration of our model, we were only able to achieve a 23.01\% BLEU score, which means that our model has low-quality output.
-
 This is caused by several factors, from the difficulty to preprocess Tibetan sentences to the lack of Tibetan-English data freely available online and our overall knowledge of the Tibetan language.
@@
 
-\subsection{RELATED WORK}
+\subsection{Related Work}
 
@@
 The non-profit 84000 provided 80 English to Tibetan parallel texts on their website \cite{c8}.
-All 80 of these texts were expert translations of Indo-Tibetan Buddhist texts.
+All 80 of these texts were expert translations of Indo-Tibetan Buddhist texts originally written in Tibetan.
 For our model we used all 80 of these provided texts.
@@
 Two such .txt file were made for each of the 80 .tmx files, a Tibetan and an English .txt file for each.
-After all 160 .txt files were generated, two single large .txt file was generated by concatenating all the Tibetan .txt files and all the English .txt files.
+After all 160 .txt files were generated, two single large .txt file were generated by concatenating all the Tibetan .txt files and all the English .txt files.
 These last two .txt files are what was fed into the model.
@@
 \includegraphics[scale=.4]{Transformer.png}
-\caption{\\Figure 1: Overview of a Transformer model \cite{11}}
+\caption{\\Figure 1: Overview of a Transformer model \cite{c11}}
 \\
@@
 Masking hides tokens from the model by replacing the tokens with a <mask> token.
-Tibetan tokens are not separated in the same way that latin languages separate their tokens.
+Tibetan tokens are not separated in the same way that Latin languages separate their tokens.
 For example,
@@
 We manually tuned the following hyperparameters: dropout, label smoothing, rnn size, the number of transformer layers and the number of heads of the attention model.
-The final configuration used can be found in the appendix.
 \par
@@
 \par
-Our six layer transformer model with switch out was unsuccessful at translating religious tibetan to english.
-The translated output does not match closely enough in either pure accuracy or in the meaning of the passage to be useful in translating religious tibetan texts to english.
+Our six layer transformer model with switch out was unsuccessful at translating religious Tibetan to English.
+The translated output does not match closely enough in either pure accuracy or in the meaning of the passage to be useful in translating religious Tibetan texts to English.
 \par
@@
 The results demonstrate the flaw in the Transformer model using the current parallel Tibetan to English corpuses.
-Each token in the Tibetan sentence corresponds to an entire phrase in english rather than a separate word.
+Each token in the Tibetan sentence corresponds to an entire phrase in English rather than a separate word.
 This complicates the translation, and the model learns parts of the phrases quite well.
 The validation accuracy is relatively high because the transformer model is good at learning parts of these phrases but it is not good at learning entire sentences or stringing together sentences in context which contributes to the low BLEU score.
-Overall, the model is capable of learning specific phrases well, but fails at the broad task of creating coherent translations from religious tibetan texts to english.
+Overall, the model is capable of learning specific phrases well, but fails at the broad task of creating coherent translations from religious Tibetan texts to English.
+\par
+We are scheduled to discuss these results with our native Tibetan speaker, Palden Gyal, but unfortunately did get to meet in time for this publication.
 
@@
 Considering our use of word segmentation for preprocessing our data (as opposed to character segmentation), our results are comparable with, albeit lower than, Tibetan-Chinese models.
-Lai, Zhao, and Bao \cite{2} ran several different Tibetan-Chinese models on the same data using two different preprocessing methodologies, character segmentation and word segmentation.
+Lai, Zhao, and Bao \cite{c2} ran several different Tibetan-Chinese models on the same data using two different preprocessing methodologies, character segmentation and word segmentation.
 Their word segmented Transformer model averaged a higher BLEU score than ours, scoring 38.44\% while our model scored 23.01\%.
@@
 
-Unlike English but similar to Chinese, Tibetan words are not clearly separated by spaces, rather, characters that represent words are strung together to make a sentence \cite{5}.
+Unlike English but similar to Chinese, Tibetan words are not clearly separated by spaces, rather, characters that represent words are strung together to make a sentence \cite{c5}.
 This makes preprocessing the data and mapping Tibetan words to their English counterparts a difficult task and it makes preprocessing on characters impossible.
@@
 On top of this, a Chinese-Tibetan model might perform overall better than a Tibetan-English one since Chinese and Tibetan are part of the same language family, and it is likely that characters may map to one another and the same preprocessing techniques can be used in both languages.
-Therefore, while comparing our model to Lai, Zhao, and Bao’s \cite{2} model may be helpful to give a general baseline on how other Tibetan translators are performing, it should not be considered where our model must perform.
+Therefore, while comparing our model to Lai, Zhao, and Bao’s \cite{c2} model may be helpful to give a general baseline on how other Tibetan translators are performing, it should not be considered where our model must perform.
 
@@
 Instead of using the translated Buddhist sutras to train our model, we could use a more general Tibetan-English corpus to train our model and then use transfer learning to fine-tune our model to religious texts.
-However, at this current time, the translated sutras provided by 84000 \cite{8} are some of the only parallel translations available on the internet.
+However, at this current time, the translated sutras provided by 84000 \cite{c8} are some of the only parallel translations available on the internet.
 
@@
 We could build an Encoder-Decoder structure based on Back Translation which would help deal with the lack of Tibetan-English translations.
-In this model, we would mix Tibetan and English sentences together to train the model \cite{10}.
+In this model, we would mix Tibetan and English sentences together to train the model \cite{c10}.
 This would help to increase the size of our training data since it would no longer just be Tibetan sentences.
@@
 
-(Chris)
 \begin{thebibliography}{99}

### 2021-04-16 paper/main.tex
Update main.tex
Last commit. Added better conclusion

@@
 Tibetan religions, in particular Buddhism, have a long tradition of translation of texts.
-While this work was not created within the context of theses traditions, it was certainly inspired by them and we hope it inspires more such translations.
+While this work was not created within the context of theses traditions, it was certainly inspired by them and we hope it will increase accessibility to Tibetan religious texts and inspire others do undergo similar projects.
 
-We have trained our models in PyTorch and have leveraged the BART architecture in the HuggingFrance transformers library to make the training and architecture more straightforward.
+We have created our Tibetan to English machine translation models in PyTorch and have leveraged the BART architecture in the HuggingFrance transformers library to make the training and architecture more straightforward.
 By creating these models, we hope to motivate scholars to study not only these texts but Tibetan as a whole.
@@
 This is caused by several factors, from the difficulty to preprocess Tibetan sentences to the lack of Tibetan-English data freely available online and our overall knowledge of the Tibetan language.
-While it may seem that we were unable to fully accomplish our lofty goal of creating a Tibetan-English translator it’s important to note that we are treading in nearly uncharted territory.
+While it may seem that we were unable to fully accomplish our lofty goal of creating a Tibetan-English translator it is important to note that we are treading in nearly uncharted territory.
 We hope that our endeavor may help future projects in this field and provide insight into the current state of Tibetan-English machine translation.
@@
 
-As a whole, not much work has been done in the area of Machine Translation or even Natural Language Processing as a whole for the Tibetan Language.
-There are a few papers exist which concern morphological segmentation and POS tagging for the Tibetan language \cite{c3} \cite{c4}.
-In recent years there has considerable interest in Chinese to Tibetan Machine translation \cite{c1} \cite{c2}.
+As a whole, not much work has been done in the area of machine translation (MT) or even Natural Language Processing (NLP) as a whole for the Tibetan language.
+There are a few papers exist which concern morphological segmentation and parts of speech (POS) tagging for the Tibetan language \cite{c3} \cite{c4}.
+In recent years there has considerable interest in Chinese to Tibetan machine translation \cite{c1} \cite{c2}.
 For example, “A Chinese to Tibetan Machine Translation System with Multiple Translating Strategies” discusses an approach of machine translation from Tibetan to Chinese.
@@
 However, Neural Network MT wasn’t prevalent at that time of its authorship.
-Nonetheless, the paper concludes, given this data (after processing), that the Statistical Based MT (implemented with conditional random fields) is indeed the best performing option.
+Nonetheless, the paper concludes, given this data (after processing), that the Statistical Based MT,implemented with conditional random fields, is indeed the best performing option.
 This is perhaps to be expected.
@@
 
-No such papers exist for Tibetan to Enlgish Machine transltions.
-In fact, to our knowledge this is the first such Tibetan to English Machine translation system developed.
-Nonetheless, the findings from the Chinese to Tibetan machine translation have given us a starting point for our investitagion.
+Unfortunately, no such papers exist for Tibetan to English machine translation.
+In fact, to our knowledge this is the first such Tibetan to English machine translation system developed.
+Nonetheless, the findings from the Chinese to Tibetan machine translation have given us a starting point for our investigation.
 
@@
 For our model we used all 80 of these provided texts.
+84000 is still actively creating more translated texts, so our model was only trained on to their most recent release which from the time of this papers' publication was March 2021.
 
@@
 That is, a newline character was used to distinguish two separate lines of translation, but the same line number in each file corresponds to the exact translation provided by 84000.
-Therefore, the total line numbers in the two .txt files were necessarily the same since one line in one file corresponding to exactly the same line of translated text in the other file.
+Therefore, the total line numbers in the two .txt files were necessarily the same since one line in one file corresponds to exactly the same line of translated text in the other file.
 Two such .txt file were made for each of the 80 .tmx files, a Tibetan and an English .txt file for each.
-After all 160 .txt files were generated, two single large .txt file were generated by concatenating all the Tibetan .txt files and all the English .txt files.
-These last two .txt files are what was fed into the model.
+After all 160 .txt files were generated, two large .txt file were generated by concatenating all the Tibetan .txt files and all the English .txt files.
+These last two large .txt files are what was fed into the model.
+The first few lines of our two large .txt files are shown below in figure 1.
 
-It was not strictly necessary to separate the creation of the large .txt files into the smaller components.
+Note in the the figure, the left hand side represents the source language (Tibetan) expressed in the Tibetan Unicode standard.
+On the right is the expert English translation provided by 84000.
+There is a one to one mapping between lines.
+This is, the Tibetan text on line one exactly corresponds to the English on line one.
+Of course, these are two separate files, and this picture simply reflects the view of the first few lines as would be seen if opened in a simple text editor.
+
+Note that it was not strictly necessary to separate the creation of the large .txt files into the smaller components.
 However, giving the limited data we had this method proved useful to see how much individual texts contributed to the model.
+
+\begin{figure*}
+  \includegraphics[width=\textwidth, height=6cm]{Screen Shot 2021-04-16 at 6.40.00 PM.png}
+  \caption{The first few lines of our fully concatenated parallel corpus}
+\end{figure*}
 
@@
 \includegraphics[scale=.4]{Transformer.png}
-\caption{\\Figure 1: Overview of a Transformer model \cite{c11}}
+\caption{\\Fig 2. Overview of a Transformer model \cite{c11}}
 \\
@@
 This phenomenon is repeated throughout the results.
-Keyphrases from the Tibetan sentence are interpreted wrongly in the same way, and then the model learns to be confident about its predictions which lowers the perplexity.
-\par
+Some examples of a few of these cases are shown in Figure 3.
+
+\newline \,\,
+\includegraphics[scale=.36]{Screen Shot 2021-04-16 at 6.54.34 PM.png}
+\caption{\\Fig 3. Example of problematic predictions }
+\newline \,\,
+
+Keyphrases from the Tibetan sentence are interpreted wrongly in the same way, and then the model learns to be confident about its predictions which lowers the perplexity. \par
 The results demonstrate the flaw in the Transformer model using the current parallel Tibetan to English corpuses.
@@
 \par
-We are scheduled to discuss these results with our native Tibetan speaker, Palden Gyal, but unfortunately did get to meet in time for this publication.
+We are scheduled to discuss these results and possible workarounds to our problems with our native Tibetan speaker, Palden Gyal, but unfortunately did get to meet in time for this publication.
 
@@
 
-As mentioned in the beginning, developing a machine learning model for Tibetan Buddhist texts is a challenging case study primarily due to the extremely low resources available.
-We overcame many obstacles while training our model and even employed the usefulness of native speakers along with publicly available resources, but our 6 layer transformer model with switch out was unsuccessful at translating religious Tibetan to English.
+As mentioned in our introduction, developing a machine learning model for Tibetan Buddhist texts is a challenging task primarily due to the extremely low resources available.
+We overcame many obstacles while training our model  while following a well defined ethical approach and using only  publicly available resources.
+However, our 6 layer transformer model with switch out was unsuccessful at translating religious Tibetan to English.
 The translated output does not match closely enough in either pure accuracy or in the meaning of the passage to be useful in translating religious Tibetan texts to English, as expanded upon in the results section.
+\par
 Additionally, we explored the ethical considerations of translating such texts and deeply reflected on those, amongst our procedure and findings.
+We faced an ethical dilemma in this part of our research project as to whether certain religious texts should or should not be translated given their sacredness.
+Another hurdle we overcame was employing switchout over masking as masking tends to hide tokens.
+We also realized that initially the result for perplexity seemed to demonstrate a strong model but further investigation brought forth that lower perplexity was actually a sign of accuracy.
+To be able to do a real world sense check and valuable input, we have scheduled a meeting with a native Tibetan speaker, through which we hope to be able to further enhance our model and increase accuracy of the same.
+\par
+To conclude, while our work is not successful, using transformer models might show promise if new approaches can find a way to add more training data or stop repeated tokens along with insights that may come from completing primary diligence in the form of conversations with native speakers.
 
@@
 The authors would like to thank Palden Gyal for volunteering to read our translated texts.
-We would also like to thank Smara Muresan, Isabelle Zuagg, and Jonathan Reeve for mentoring this research.
+We would also like to thank Smara Muresan, Isabelle Zaugg, and Jonathan Reeve for mentoring this research.
+We would like to thank Professor Thomas Yarnall and Professor Paul Hackett for providing additional advice.
 Lastly, we would like to thank the 84000 project for providing open source parallel Tibetan corpora.
